---
title: "Homework 2"
author: "ruby"
date: "2025-09-26"
output: github_document
---

## Problem 1

Install libraries
```{r}
library(tidyverse)
library(readxl)
```

Upload and clean pols-month.csv dataset
```{r}
pols_df = 
  read_csv("fivethirtyeight_datasets/pols-month.csv") |> 
  janitor::clean_names() |> 
  separate(mon,into = c("year", "month", "day"), sep = "-", convert = TRUE) |> 
  mutate(month = month.name[month], president = prez_dem, prez_gop ) |>   
  select(-day, -prez_dem, - prez_gop)
```


upload and clean snp.csv dataset
```{r}
snp_df = 
  read_csv("fivethirtyeight_datasets/snp.csv") |> 
  janitor::clean_names() |> 
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) |>
  select(year, month, everything()) |> 
  mutate(month = month.name[month], 
         year = ifelse(year < 30, year + 2000, year + 1900)
         )
```

upload and clean unemployment.csv 
```{r}
unemploy_df = 
  read_csv("fivethirtyeight_datasets/unemployment.csv") |> 
  janitor::clean_names() |> 
  pivot_longer(jan:dec, 
               names_to = "month", 
               values_to = "unemployment %") |> 
  mutate(month = month.name[match(tolower(month), tolower(month.abb))])
```

Join the datasets by merging snp into pols, and merging unemployment into the result. (NEED TO COME BACK HERE TO SEE WHAT KIND OF JOIN TO USE)
```{r}
prob1_df =
  left_join(snp_df, pols_df, by = "year")
```

## Problem 2

import in mr trash wheel dataset 
```{r}
mrtrashwheel_df = 
  read_excel("202409 Trash Wheel Collection Data.xlsx", na = c(".", "NA", ""))
```

specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data
round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)